Predicting the Capture Rate of Pokemon

### 2022-12-10 Final_Draft.Rmd
commit

```{r}
# Adding in 5 more variables indicating whether a Pokemon has each of the top 5 abilities or not
pokemon <- pokemon %>% 
  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>%
  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))

# Creating a final dataset csv
write_csv(pokemon, "pokemon_final.csv")
```

@@

```{r, message = FALSE}
# Reading in the data
pokemon_data <- read_csv("pokemon_final.csv")

@@

# RIDGE REGRESSION
write_rds(ridge_tune, file = "ridge.rds")

# LASSO REGRESSION
write_rds(lasso_tune, file = "lasso.rds")

# POLYNOMIAL REGRESSION
write_rds(poly_tuned, file = "poly.rds")

# K NEAREST NEIGHBORS
write_rds(knn_tune, file = "knn.rds")

# ELASTIC NET
write_rds(elastic_tune, file = "elastic.rds")

# RANDOM FOREST
write_rds(rf_tune_res, file = "rf.rds")

# BOOSTED TREES
write_rds(boosted_tune_res, file = "boosted.rds")
```

@@

# RIDGE REGRESSION
ridge_tuned <- read_rds(file = "ridge.rds")

# LASSO REGRESSION
lasso_tuned <- read_rds(file = "lasso.rds")

# POLYNOMIAL REGRESSION
poly_tuned <- read_rds(file = "poly.rds")

# K NEAREST NEIGHBORS
knn_tuned <- read_rds(file = "knn.rds")

# ELASTIC NET
elastic_tuned <- read_rds(file = "elastic.rds")

# RANDOM FOREST
rf_tuned <- read_rds(file = "rf.rds")

# BOOSTED TREES
boosted_tuned <- read_rds(file = "boosted.rds")
```

@@

# LINEAR REGRESSION 
# No need to select the best fit as linear regression had no tuning -> we just fit the one linear regression model we fit
lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
write_rds(lm_fit, file = "lm_final.rds")

# RIDGE REGRESSION
best_penalty <- select_best(ridge_tune, metric = 'rmse')
ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
write_rds(ridge_final_fit, file = "ridge_final.rds")

# LASSO REGRESSION
best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
write_rds(lasso_final_fit, file = "lasso_final.rds")

# POLYNOMIAL REGRESSION
best_poly<- select_best(poly_tuned, metric = "rmse")
poly_final <- finalize_workflow(poly_wf, best_poly)
poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
write_rds(poly_final_fit, file = "poly_final.rds")

# K NEAREST NEIGHBORS
best_knn <- select_best(knn_tune, metric = "rmse")
knn_final <- finalize_workflow(knn_workflow, best_knn)
knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
write_rds(knn_final_fit, file = "knn_final.rds")

# ELASTIC NET
best_elastic <- select_best(elastic_tune, metric = "rmse")
elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
write_rds(elastic_final_fit, file = "elastic_final.rds")

# RANDOM FOREST
best_rf <- select_best(rf_tuned)
rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
write_rds(rf_final_fit, file = "rf_final.rds")

# BOOSTED TREES
best_boosted <- select_best(boosted_tuned)
boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
write_rds(boosted_final_fit, file = "boosted_final.rds")
```

@@

# LINEAR REGRESSION 
lm_final_fit <- read_rds(file = "lm_final.rds")
lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)

# RIDGE REGRESSION
ridge_final_fit <- read_rds(file = "ridge_final.rds")
ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)

# LASSO REGRESSION
lasso_final_fit <- read_rds(file = "lasso_final.rds")
lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)

# POLYNOMIAL REGRESSION
poly_final_fit <- read_rds(file = "poly_final.rds")
poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)

# K NEAREST NEIGHBORS
knn_final_fit <- read_rds(file = "knn_final.rds")
knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)

# ELASTIC NET
elastic_final_fit <- read_rds(file = "elastic_final.rds")
elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)

# RANDOM FOREST
rf_final_fit <- read_rds(file = "rf_final.rds")
rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)

# BOOSTED TREES
boosted_final_fit <- read_rds(file = "boosted_final.rds")
boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
```

@@

For the random forest,

@@

```{r, eval = FALSE}
# Fitting to the training data
best_rf_train <- select_best(rf_tuned)
rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit_train <- fit(rf_final_workflow, data = pokemon_train)

write_rds(rf_final_fit, file = "rf_final_train.rds")
```

```{r}
rf_final_fit_train <- read_rds(file = "rf_final_train.rds")

# VIP plot
rf_final_fit_train %>% 
  extract_fit_engine() %>% 
  vip()
```

```{r}
# Indicating the desired metric
pokemon_metric <- metric_set(rmse)

# Creating the predicted vs. actual value tibble
pokemon_tibble <- predict(rf_final_fit, new_data = pokemon_test %>% select(-capture_rate))
pokemon_tibble <- bind_cols(pokemon_tibble, pokemon_test %>% select(capture_rate))

# Collecting the rmse of the model on the testing data
pokemon_tibble_metrics <- pokemon_metric(pokemon_tibble, truth = capture_rate, estimate = .pred)
pokemon_tibble_metrics
```

>>>>

```{r, eval=FALSE}
# Adding in 5 more variables indicating whether a Pokemon has each of the top 5 abilities or not
pokemon <- pokemon %>% 
  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>%
  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))

# Creating a final dataset csv
write_csv(pokemon, "data/pokemon_final.csv")
```

@@

```{r, message = FALSE}
# Reading in the data
pokemon_data <- read_csv("data/pokemon_final.csv")

@@

# RIDGE REGRESSION
write_rds(ridge_tune, file = "data/tuned_models/ridge.rds")

# LASSO REGRESSION
write_rds(lasso_tune, file = "data/tuned_models/lasso.rds")

# POLYNOMIAL REGRESSION
write_rds(poly_tuned, file = "data/tuned_models/poly.rds")

# K NEAREST NEIGHBORS
write_rds(knn_tune, file = "data/tuned_models/knn.rds")

# ELASTIC NET
write_rds(elastic_tune, file = "data/tuned_models/elastic.rds")

# RANDOM FOREST
write_rds(rf_tune_res, file = "data/tuned_models/rf.rds")

# BOOSTED TREES
write_rds(boosted_tune_res, file = "data/tuned_models/boosted.rds")
```

@@

# RIDGE REGRESSION
ridge_tuned <- read_rds(file = "data/tuned_models/ridge.rds")

# LASSO REGRESSION
lasso_tuned <- read_rds(file = "data/tuned_models/lasso.rds")

# POLYNOMIAL REGRESSION
poly_tuned <- read_rds(file = "data/tuned_models/poly.rds")

# K NEAREST NEIGHBORS
knn_tuned <- read_rds(file = "data/tuned_models/knn.rds")

# ELASTIC NET
elastic_tuned <- read_rds(file = "data/tuned_models/elastic.rds")

# RANDOM FOREST
rf_tuned <- read_rds(file = "data/tuned_models/rf.rds")

# BOOSTED TREES
boosted_tuned <- read_rds(file = "data/tuned_models/boosted.rds")
```

@@

# LINEAR REGRESSION 
# No need to select the best fit as linear regression had no tuning -> we just fit the one linear regression model we fit
lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
write_rds(lm_fit, file = "data/final_models/lm_final.rds")

# RIDGE REGRESSION
best_penalty <- select_best(ridge_tune, metric = 'rmse')
ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
write_rds(ridge_final_fit, file = "data/final_models/ridge_final.rds")

# LASSO REGRESSION
best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
write_rds(lasso_final_fit, file = "data/final_models/lasso_final.rds")

# POLYNOMIAL REGRESSION
best_poly<- select_best(poly_tuned, metric = "rmse")
poly_final <- finalize_workflow(poly_wf, best_poly)
poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
write_rds(poly_final_fit, file = "data/final_models/poly_final.rds")

# K NEAREST NEIGHBORS
best_knn <- select_best(knn_tune, metric = "rmse")
knn_final <- finalize_workflow(knn_workflow, best_knn)
knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
write_rds(knn_final_fit, file = "data/final_models/knn_final.rds")

# ELASTIC NET
best_elastic <- select_best(elastic_tune, metric = "rmse")
elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
write_rds(elastic_final_fit, file = "data/final_models/elastic_final.rds")

# RANDOM FOREST
best_rf <- select_best(rf_tuned)
rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
write_rds(rf_final_fit, file = "data/final_models/rf_final.rds")

# BOOSTED TREES
best_boosted <- select_best(boosted_tuned)
boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
write_rds(boosted_final_fit, file = "data/final_models/boosted_final.rds")
```

@@

# LINEAR REGRESSION 
lm_final_fit <- read_rds(file = "data/final_models/lm_final.rds")
lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)

# RIDGE REGRESSION
ridge_final_fit <- read_rds(file = "data/final_models/ridge_final.rds")
ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)

# LASSO REGRESSION
lasso_final_fit <- read_rds(file = "data/final_models/lasso_final.rds")
lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)

# POLYNOMIAL REGRESSION
poly_final_fit <- read_rds(file = "data/final_models/poly_final.rds")
poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)

# K NEAREST NEIGHBORS
knn_final_fit <- read_rds(file = "data/final_models/knn_final.rds")
knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)

# ELASTIC NET
elastic_final_fit <- read_rds(file = "data/final_models/elastic_final.rds")
elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)

# RANDOM FOREST
rf_final_fit <- read_rds(file = "data/final_models/rf_final.rds")
rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)

# BOOSTED TREES
boosted_final_fit <- read_rds(file = "data/final_models/boosted_final.rds")
boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
```

@@

For the random forest, we tuned the the minimal node size, the number of randomly selected predictors, and the number of trees. The number of trees does not seem to have much effect on the performance of the model. The minimal node size does not appear to have a huge effect on the performance, but the smaller values of `min_n` appear to have slightly lower RMSE values (at least judging by the starting point). The optimal node size appears to be at 5 (which lines up with the aforementioned observation). The number of predictors appears to have a greatest effect on performance. From the plots, it appears that, overall, a greater number of predictors renders a better performance. Actually, it seems that including all 14 predictors gives the lowest RMSE here. Based on the RMSE values in the plots, this model definitely performs the best!

@@

## Performance on the Folds

So, the random forest performed the best out of all 8 of our models. But which tuned parameters were chosen as the best random forest model?

```{r}
# Collecting metrics on the tuned random forest and slicing the lowest RMSE
rf_tuned %>% 
  collect_metrics() %>% 
  arrange(mean) %>% 
  slice(513)
```

Random forest #40 with all 14 predictors, 657 trees, and a minimal node size of 5 performed the best with an RMSE of 43.79638!

## Fitting to Training Data

Now, we will take that best model from the tuned random forest and fit it to the training data. This will train that random forest one more time on the entire training data set.

```{r, eval = FALSE, warning = FALSE}
# Fitting to the training data
best_rf_train <- select_best(rf_tuned)
rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit_train <- fit(rf_final_workflow, data = pokemon_train)

# Save
write_rds(rf_final_fit, file = "data/rf_final/rf_final_train.rds")
```

```{r}
# Reading in the random forest fit on the training data
rf_final_fit_train <- read_rds(file = "data/rf_final/rf_final_train.rds")

# Collecting the RMSE of the random forest fit on the training data
rf_final_fit_train %>% 
  collect_metrics() %>% 
  slice(1)
```

The chosen random forest model had an RMSE of 43.95341 on the training data. This is quite close to the RMSE the model had on the folds.

## Testing the Model

Now, it's time to test our random forest model to see how it performs on data that it has not been trained on at all: the testing data set.

```{r, eval = FALSE}
# Creating the predicted vs. actual value tibble
pokemon_tibble <- predict(rf_final_fit_train, new_data = pokemon_test %>% select(-capture_rate))
pokemon_tibble <- bind_cols(pokemon_tibble, pokemon_test %>% select(capture_rate))

# Save
write_rds(pokemon_tibble, file = "data/rf_final/final_model.rds")
```

```{r}
# Load in final model
pokemon_tibble <- read_rds(file = "data/rf_final/final_model.rds")

# Indicating the desired metric
pokemon_metric <- metric_set(rmse)

# Collecting the rmse of the model on the testing data
pokemon_tibble_metrics <- pokemon_metric(pokemon_tibble, truth = capture_rate, estimate = .pred)
pokemon_tibble_metrics

```

Our random forest actually performed slightly better on the testing set than on the training set and folds with an RMSE of 42.42378!
